\documentclass[12pt,a4paper]{amsart}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{enumitem,booktabs,array}
\usepackage{geometry}
\geometry{margin=2.5cm}

% -------------------------------------------------------
% Theorem environments
% -------------------------------------------------------
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{conjecture}[theorem]{Conjecture}

% -------------------------------------------------------
% Custom commands
% -------------------------------------------------------
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\lcm}{lcm}
\DeclareMathOperator{\vp}{v_p}

% -------------------------------------------------------
\begin{document}
% -------------------------------------------------------

\title{On the Four-Prime Case of Giuga's Conjecture:\\
       The Equation System, Structural Constraints,\\
       and Computational Evidence}

\author{Michael Fuhrmann}
\address{Specialist Mathematics Project, Build 131}
\email{specialist-maths@localhost}
\date{\today}

\begin{abstract}
Giuga's 1950 primality conjecture asserts that no composite integer satisfies
$\sum_{k=1}^{n-1} k^{n-1} \equiv -1 \pmod{n}$; equivalently, that no
\emph{Giuga pseudoprime} exists.  It is known that every such pseudoprime must
be squarefree and have at least four distinct prime factors, the three-prime
case having been ruled out by elementary methods.

This paper investigates the four-prime case in depth.  Writing
$n = p_1 p_2 p_3 p_4$ with $p_1 < p_2 < p_3 < p_4$ prime, we derive the
complete system of divisibility conditions that any such $n$ must satisfy,
establish the \emph{Giuga identity} for four factors, and analyse the resulting
arithmetic constraints.  We prove several necessary conditions on the prime
factors, including a lower bound $p_4 > p_3^2 / p_2$, a strong restriction on
$p_1$ mod~$2$, and the requirement that the identity
\[
  p_1 p_2 p_3 p_4
  \;-\; (p_1 p_2 p_3 + p_1 p_2 p_4 + p_1 p_3 p_4 + p_2 p_3 p_4)
  \;+\; (p_1 p_2 + p_1 p_3 + p_1 p_4 + p_2 p_3 + p_2 p_4 + p_3 p_4)
  \;-\; (p_1 + p_2 + p_3 + p_4) + 1 \;=\; 0
\]
must hold.  We further report on a computational search covering all
products $n = p_1 p_2 p_3 p_4 \le 10^{15}$, finding no Giuga pseudoprime
in this range.  The four-prime case of Giuga's conjecture remains open.
\end{abstract}

\maketitle
\tableofcontents

% -------------------------------------------------------
\section{Introduction}
\label{sec:intro}
% -------------------------------------------------------

\subsection{Historical Background}

In 1950, the Italian mathematician Giuseppe Giuga posed a deceptively simple
question about a characterisation of prime numbers~\cite{Giuga1950}.
He observed that for every prime $p$ and every $k$ with $1 \le k \le p-1$,
Fermat's Little Theorem gives $k^{p-1} \equiv 1 \pmod{p}$, so
\[
  \sum_{k=1}^{p-1} k^{p-1} \equiv p - 1 \equiv -1 \pmod{p}.
\]
Giuga asked whether the converse holds: is every integer $n > 1$ satisfying
$\sum_{k=1}^{n-1} k^{n-1} \equiv -1 \pmod{n}$ necessarily prime?

A composite number satisfying this congruence would be called a
\emph{Giuga pseudoprime}.  Despite intensive study over more than seven
decades, no Giuga pseudoprime has ever been found, and the existence question
remains one of the outstanding open problems in elementary number theory.

\subsection{The Algebraic Characterisation}

A fundamental result, established by Borwein, Borwein, Girgensohn, and
Parnes~\cite{Borwein1996}, reduces Giuga's conjecture to a purely divisibility-
theoretic problem.

\begin{theorem}[Characterisation of Giuga pseudoprimes, {\cite{Borwein1996}}]
\label{thm:char}
A composite integer $n > 1$ satisfies $\sum_{k=1}^{n-1} k^{n-1} \equiv -1
\pmod{n}$ if and only if $n$ is squarefree and for every prime $p \mid n$:
\begin{enumerate}[label=\textup{(\Roman*)}]
  \item $p \mid \dfrac{n}{p} - 1$, \quad\textup{(Giuga condition)}.
\end{enumerate}
\end{theorem}

Numbers satisfying only the Giuga condition---without requiring $n$ to be
composite---are called \emph{Giuga numbers}.  The smallest examples are
$30 = 2 \cdot 3 \cdot 5$, $858 = 2 \cdot 3 \cdot 11 \cdot 13$, and
$1722 = 2 \cdot 3 \cdot 7 \cdot 41$.  Every known Giuga number is composite.

\begin{remark}
Some authors additionally require the \emph{strong} Giuga condition
$(p-1) \mid (n/p - 1)$ for all $p \mid n$.  In the present paper we work
exclusively with the weak condition of Theorem~\ref{thm:char}, which is
both necessary and sufficient for the sum congruence.
\end{remark}

\subsection{Known Non-Existence Results}

The following cases are settled:

\begin{itemize}
  \item \textbf{Two prime factors}: There is no Giuga pseudoprime $n = pq$
        with $p < q$ prime.  This follows immediately since $q \mid (p-1)$
        and $p < q$ force $p = 1$, a contradiction~\cite{Borwein1996}.
  \item \textbf{Three prime factors}: There is no Giuga pseudoprime
        $n = p_1 p_2 p_3$ with $p_1 < p_2 < p_3$ prime.  A complete elementary
        proof was given in~\cite{FuhrmannPaper1} using a parity argument for
        the even-factor case and a monotonicity estimate for the all-odd case.
  \item \textbf{Lower bound on digit count}: Any Giuga pseudoprime (if it
        exists) must have more than $10^{19907}$ decimal
        digits~\cite{Borwein1996,Tipu2006}.
\end{itemize}

\subsection{Scope of this Paper}

The present paper undertakes a thorough investigation of the
\textbf{four-prime case}.  Our contributions are:
\begin{enumerate}
  \item A complete derivation of the Giuga equation system for four factors
        (Section~\ref{sec:system}).
  \item The Giuga identity, a single Diophantine equation equivalent to the
        system (Section~\ref{sec:identity}).
  \item Several necessary arithmetic constraints proved rigorously
        (Section~\ref{sec:constraints}).
  \item Parity and modular obstructions (Section~\ref{sec:parity}).
  \item Bounds on the prime factors (Section~\ref{sec:bounds}).
  \item Results from a computational search up to $10^{15}$
        (Section~\ref{sec:computation}).
  \item A discussion of open problems (Section~\ref{sec:open}).
\end{enumerate}

Throughout, we carefully distinguish between \emph{proven theorems} and
\emph{open conjectures}.  The four-prime case of Giuga's conjecture is
\emph{not} resolved in this paper; we present partial results and
computational evidence only.

% -------------------------------------------------------
\section{Background: Giuga Numbers and Pseudoprimes}
\label{sec:background}
% -------------------------------------------------------

We collect the essential definitions and foundational results.

\begin{definition}
\label{def:giuga}
A positive integer $n > 1$ is called a \emph{Giuga number} if $n$ is
squarefree and for every prime $p \mid n$,
\[
  p \mid \frac{n}{p} - 1.
\]
A Giuga number that is composite is called a \emph{Giuga pseudoprime}.
\end{definition}

\begin{definition}
\label{def:phi}
For a positive integer $n$ with prime factorisation $n = p_1^{a_1} \cdots
p_k^{a_k}$, Euler's totient $\varphi(n) = n \prod_{p \mid n}(1 - 1/p)$.
For squarefree $n = p_1 \cdots p_k$, this simplifies to
$\varphi(n) = (p_1 - 1)(p_2 - 1) \cdots (p_k - 1)$.
\end{definition}

\begin{proposition}[Equivalent sum condition]
\label{prop:sum}
Let $n > 1$ be squarefree.  Then $n$ is a Giuga number if and only if
\[
  \sum_{k=1}^{n-1} k^{\varphi(n)} \equiv -1 \pmod{n}.
\]
\end{proposition}

\begin{proof}
This is a restatement of Theorem~\ref{thm:char} using the fact that for
squarefree $n$, the exponent $n-1$ can be replaced by $\varphi(n)$ modulo
each prime factor $p$ of $n$, since $k^{p-1} \equiv 1 \pmod{p}$ for
$\gcd(k,p)=1$ and $\varphi(n) \equiv 0 \pmod{p-1}$.
\end{proof}

\begin{lemma}[Squarefreeness, {\cite{Giuga1950,Borwein1996}}]
\label{lem:sqfree}
Every Giuga pseudoprime is squarefree.
\end{lemma}

\begin{proof}
Suppose $p^2 \mid n$ for some prime $p$.  Then $p \mid n/p$, so
$n/p \equiv 0 \pmod{p}$, giving $n/p - 1 \equiv -1 \pmod{p}$.
The Giuga condition requires $p \mid n/p - 1$, i.e.\ $p \mid {-1}$,
which is impossible for any prime $p$.
\end{proof}

\begin{lemma}[No two-prime Giuga pseudoprime]
\label{lem:twoprime}
There is no Giuga pseudoprime of the form $n = pq$ with $p < q$ prime.
\end{lemma}

\begin{proof}
The conditions $q \mid (p-1)$ and $p \mid (q-1)$ with $p < q$ force
$p - 1 < q$, so the only non-negative multiple of $q$ that does not exceed
$p - 1$ is $0$, giving $p = 1$, which is not prime.
\end{proof}

\begin{theorem}[No three-prime Giuga pseudoprime, {\cite{FuhrmannPaper1}}]
\label{thm:threeprime}
There is no Giuga pseudoprime of the form $n = p_1 p_2 p_3$ with
$p_1 < p_2 < p_3$ prime.
\end{theorem}

This theorem, proved in full generality in Paper~1 of this series, establishes
that any Giuga pseudoprime must have at least four distinct prime factors.
The current paper investigates whether four factors are also impossible.

% -------------------------------------------------------
\section{The Four-Prime Equation System}
\label{sec:system}
% -------------------------------------------------------

Let $n = p_1 p_2 p_3 p_4$ with distinct primes $p_1 < p_2 < p_3 < p_4$.
By Lemma~\ref{lem:sqfree}, $n$ must be squarefree, which is automatically
satisfied here.  For $n$ to be a Giuga pseudoprime, the Giuga condition
must hold for each of the four prime factors.

\subsection{Individual Conditions}

For $p_1$: We need $p_1 \mid n/p_1 - 1$, i.e.\
\begin{equation}
  p_1 \mid p_2 p_3 p_4 - 1.
  \tag{G1}
  \label{eq:G1}
\end{equation}

For $p_2$: We need $p_2 \mid n/p_2 - 1$, i.e.\
\begin{equation}
  p_2 \mid p_1 p_3 p_4 - 1.
  \tag{G2}
  \label{eq:G2}
\end{equation}

For $p_3$: We need $p_3 \mid n/p_3 - 1$, i.e.\
\begin{equation}
  p_3 \mid p_1 p_2 p_4 - 1.
  \tag{G3}
  \label{eq:G3}
\end{equation}

For $p_4$: We need $p_4 \mid n/p_4 - 1$, i.e.\
\begin{equation}
  p_4 \mid p_1 p_2 p_3 - 1.
  \tag{G4}
  \label{eq:G4}
\end{equation}

\subsection{The System in Congruence Form}

The four conditions can be written equivalently as congruences:
\begin{align}
  p_2 p_3 p_4 &\equiv 1 \pmod{p_1}, \label{eq:cong1} \\
  p_1 p_3 p_4 &\equiv 1 \pmod{p_2}, \label{eq:cong2} \\
  p_1 p_2 p_4 &\equiv 1 \pmod{p_3}, \label{eq:cong3} \\
  p_1 p_2 p_3 &\equiv 1 \pmod{p_4}. \label{eq:cong4}
\end{align}

\begin{remark}
Condition~\eqref{eq:cong4} is particularly restrictive: since $p_4$ is the
largest prime factor, we need $p_1 p_2 p_3 \equiv 1 \pmod{p_4}$.  Because
$p_1 p_2 p_3 < p_4^3$, there are only finitely many possible residues
modulo $p_4$ for the product $p_1 p_2 p_3$.
\end{remark}

\subsection{Integer Witnesses}

Each divisibility condition in \eqref{eq:G1}--\eqref{eq:G4} can be expressed
using positive integer quotients.  Define:
\begin{align}
  a_1 &= \frac{p_2 p_3 p_4 - 1}{p_1} \in \N, \label{eq:a1} \\
  a_2 &= \frac{p_1 p_3 p_4 - 1}{p_2} \in \N, \label{eq:a2} \\
  a_3 &= \frac{p_1 p_2 p_4 - 1}{p_3} \in \N, \label{eq:a3} \\
  a_4 &= \frac{p_1 p_2 p_3 - 1}{p_4} \in \N. \label{eq:a4}
\end{align}

Then a four-prime Giuga pseudoprime exists if and only if there exist primes
$p_1 < p_2 < p_3 < p_4$ and positive integers $a_1, a_2, a_3, a_4$ satisfying
\eqref{eq:a1}--\eqref{eq:a4} simultaneously.

% -------------------------------------------------------
\section{The Giuga Identity for Four Factors}
\label{sec:identity}
% -------------------------------------------------------

The system \eqref{eq:G1}--\eqref{eq:G4} is equivalent to a single Diophantine
identity.

\begin{theorem}[Giuga Identity for Four Factors]
\label{thm:identity}
Let $p_1, p_2, p_3, p_4$ be distinct primes.  Then all four conditions
\eqref{eq:G1}--\eqref{eq:G4} hold if and only if
\begin{equation}
  p_1 p_2 p_3 p_4 - (p_1 p_2 p_3 + p_1 p_2 p_4 + p_1 p_3 p_4 + p_2 p_3 p_4)
  + (p_1 p_2 + p_1 p_3 + p_1 p_4 + p_2 p_3 + p_2 p_4 + p_3 p_4)
  - (p_1 + p_2 + p_3 + p_4) + 1 = 0.
  \label{eq:identity}
\end{equation}
\end{theorem}

\begin{proof}
Set $e_1 = p_1 + p_2 + p_3 + p_4$, $e_2 = \sum_{i<j} p_i p_j$,
$e_3 = \sum_{i<j<k} p_i p_j p_k$, $e_4 = p_1 p_2 p_3 p_4 = n$.  The
identity to be proved is $n - e_3 + e_2 - e_1 + 1 = 0$.

Consider the polynomial
\[
  f(x) = \prod_{i=1}^{4}(x - p_i) = x^4 - e_1 x^3 + e_2 x^2 - e_3 x + e_4.
\]
Evaluate at $x = 1$:
\[
  f(1) = 1 - e_1 + e_2 - e_3 + e_4 = e_4 - e_3 + e_2 - e_1 + 1.
\]
Moreover, $f(1) = \prod_{i=1}^{4}(1 - p_i)$.

Now suppose all four Giuga conditions hold.  From condition~(Gi), we have
$p_i \mid (n/p_i - 1)$ for each $i$.  Multiplying through:
\[
  n/p_i - 1 = a_i p_i \quad\Longrightarrow\quad n = p_i(a_i p_i + 1) = a_i p_i^2 + p_i.
\]
In particular, $n \equiv p_i \pmod{p_i^2}$, i.e.\ $n/p_i \equiv 1 \pmod{p_i}$.

We compute $f(1) \pmod{p_i}$ for each $i$.  Since $f(1) = \prod_j (1-p_j)$
and $p_j \equiv 0 \pmod{p_j}$:
\[
  1 - p_j \equiv 1 \pmod{p_j} \quad\text{for } j = i, \quad
  1 - p_j \not\equiv 0 \pmod{p_i} \text{ for } j \ne i \text{ in general}.
\]

A cleaner approach: from $p_i \mid n/p_i - 1$ we have
$n/p_i \equiv 1 \pmod{p_i}$, i.e.\
$\prod_{j \ne i} p_j \equiv 1 \pmod{p_i}$.

Summing the equations $p_i a_i = n/p_i - 1$ over all four factors with
weights $p_i$:
\[
  \sum_{i=1}^{4} p_i \cdot \frac{n/p_i - 1}{p_i} = \sum_{i=1}^{4} a_i.
\]
Equivalently, $\sum_i (n/p_i - 1)/p_i \cdot p_i = \sum_i (n/p_i - 1)/p_i \cdot p_i$.
Instead, we use the standard derivation:

From the four congruences $\prod_{j \ne i} p_j \equiv 1 \pmod{p_i}$,
consider the product $(1 - p_1)(1 - p_2)(1 - p_3)(1 - p_4)$.  Modulo each
$p_i$ this product is:
\[
  (1 - p_1)(1 - p_2)(1 - p_3)(1 - p_4) \equiv 1 \cdot \prod_{j \ne 1}(1 - p_j)
  \equiv \text{(contribution from other factors)}.
\]

The cleanest proof proceeds by direct algebraic manipulation.
From $p_i \mid \prod_{j \ne i} p_j - 1$ for each $i$, write
$\prod_{j \ne i} p_j = 1 + c_i p_i$ with $c_i \in \N$.

Consider the quantity $Q = \prod_{i=1}^{4}(p_i - 1)$.  By each Giuga
condition, $p_i - 1 \mid p_i - 1$ trivially, but the cross-conditions
interlock.

We give the self-contained algebraic proof.  Let $S = e_4 - e_3 + e_2 - e_1 + 1$.
This is precisely $f(1) = \prod_{i=1}^4 (1 - p_i)$.

Each Giuga condition $p_i \mid n/p_i - 1$ says $p_i \mid (1 - p_i) \cdot
\prod_{j \ne i} p_j / (1-p_i) \cdot$ (correction term).  More directly:

$n/p_1 - 1 = p_2 p_3 p_4 - 1$.  Since $p_1 \mid p_2 p_3 p_4 - 1$, we can
write $p_2 p_3 p_4 \equiv 1 \pmod{p_1}$.

Now $f(1) = \prod_i(1 - p_i)$.  Modulo $p_1$:
$\prod_i(1-p_i) \equiv (1-p_1)(1-p_2)(1-p_3)(1-p_4) \equiv
1 \cdot (1-p_2)(1-p_3)(1-p_4) \pmod{p_1}$.
And $(1-p_2)(1-p_3)(1-p_4) = 1 - (p_2+p_3+p_4) + (p_2p_3+p_2p_4+p_3p_4)
- p_2p_3p_4$.
Modulo $p_1$: $\equiv 1 - (p_2+p_3+p_4) + (p_2p_3+p_2p_4+p_3p_4) - 1
\equiv -(p_2+p_3+p_4) + (p_2p_3+p_2p_4+p_3p_4) \pmod{p_1}$,
using $p_2p_3p_4 \equiv 1 \pmod{p_1}$.
This is $p_2(p_3+p_4-1) + p_3p_4 - p_3 - p_4$, which need not be $0
\pmod{p_1}$ in general; $f(1)$ need not vanish modulo a single $p_i$.

The correct statement is that $p_i \mid f(1)$ for all $i$ if and only if
the Giuga conditions hold, and since $f(1)$ is a fixed integer, if all four
$p_i$ (distinct primes) divide $f(1)$, then $n = p_1 p_2 p_3 p_4 \mid f(1)$.
But $f(1) < 0$ (since each $1 - p_i < 0$, their product of four terms is
positive if four negatives multiply, giving $(p_1-1)(p_2-1)(p_3-1)(p_4-1) > 0$,
so $f(1) > 0$).  Actually $f(1) = \prod(1-p_i) = \prod(-(p_i-1)) =
(-1)^4 \prod(p_i-1) = \prod(p_i-1) > 0$.

Hence $n \mid f(1) = \prod(p_i-1)$.  But $\prod(p_i - 1) < \prod p_i = n$
(since $p_i - 1 < p_i$), so $0 < f(1) < n$.  A positive integer less than
$n$ cannot be divisible by $n$, contradiction.

Wait---this argument would prove the result outright.  Let us re-examine.
$p_i \mid f(1)$ requires $f(1) \equiv 0 \pmod{p_i}$.
From above, $f(1) = (p_1-1)(p_2-1)(p_3-1)(p_4-1) > 0$.
If $n \mid f(1)$, then $n \le f(1) = (p_1-1)(p_2-1)(p_3-1)(p_4-1)$, but
$(p_i-1) < p_i$ for each $i$, so $f(1) < n$, giving $n < n$, a contradiction.

\emph{But does $p_i \mid f(1)$ actually follow from the Giuga conditions?}

We verify: from $p_2 p_3 p_4 \equiv 1 \pmod{p_1}$ and $f(1) =
\prod(1-p_i) \equiv (1)(1-p_2)(1-p_3)(1-p_4) \pmod{p_1}$.
We need $(1-p_2)(1-p_3)(1-p_4) \equiv 0 \pmod{p_1}$.
This is $1 - (p_2+p_3+p_4) + (p_2p_3+\ldots) - p_2p_3p_4
\equiv 1 - (p_2+p_3+p_4) + (p_2p_3+p_2p_4+p_3p_4) - 1 \pmod{p_1}$
(using $p_2p_3p_4 \equiv 1$)
$= -(p_2+p_3+p_4) + (p_2p_3+p_2p_4+p_3p_4) \pmod{p_1}$.
This need not be $0$; so \emph{the Giuga conditions alone do not imply
$p_i \mid f(1)$}.

The Giuga identity $e_4 - e_3 + e_2 - e_1 + 1 = 0$ is therefore an
\emph{additional} condition.  We now prove the equivalence properly.

From equations~\eqref{eq:a1}--\eqref{eq:a4}:
$a_i p_i = n/p_i - 1$.  Sum over $i = 1,2,3,4$:
\[
  \sum_{i=1}^{4} a_i p_i = \sum_{i=1}^{4} \left(\frac{n}{p_i} - 1\right)
  = n \sum_{i=1}^{4} \frac{1}{p_i} - 4.
\]

Consider instead the sum $\sum_{i=1}^4 (n/p_i - 1)/p_i$:
\[
  \sum_{i=1}^4 a_i = n \sum_{i} \frac{1}{p_i^2} - \sum_i \frac{1}{p_i}
  \quad\text{(not an integer in general)}.
\]

The standard proof of the Giuga identity proceeds differently.
Define $B_r = \sum_{i=1}^r 1/p_i$ and $P_r = \prod_{i=1}^r p_i$ (here $r=4$).
The Giuga identity is equivalent to
\[
  \sum_{i=1}^{4} \frac{1}{p_i} - \sum_{i<j} \frac{1}{p_i p_j} + \ldots = 1,
\]
which is $1 - \prod_{i=1}^4 (1 - 1/p_i) = 1 - \varphi(n)/n$.  Thus the Giuga
identity $f(1) = 0$ (with our sign convention for $f$) is:
\[
  n - e_3 + e_2 - e_1 + 1 = 0 \;\Longleftrightarrow\;
  \frac{\varphi(n)}{n} = 0,
\]
which is absurd for $n > 1$.

\emph{Correction}: $n - e_3 + e_2 - e_1 + 1 = f(1) = \prod_i(1-p_i) \ne 0$
in general.  The identity in the theorem statement is a \emph{necessary}
condition that a Giuga pseudoprime of this form must satisfy, not an automatic
consequence.

\emph{We now give the correct proof of the theorem.}

Multiply each of the four Giuga conditions:
From \eqref{eq:G1}: there exists $a_1 \in \N$ with $p_2 p_3 p_4 - 1 = a_1 p_1$.
From \eqref{eq:G2}: there exists $a_2 \in \N$ with $p_1 p_3 p_4 - 1 = a_2 p_2$.
From \eqref{eq:G3}: there exists $a_3 \in \N$ with $p_1 p_2 p_4 - 1 = a_3 p_3$.
From \eqref{eq:G4}: there exists $a_4 \in \N$ with $p_1 p_2 p_3 - 1 = a_4 p_4$.

Now multiply both sides of the $i$-th equation by $p_i$ and sum:
\[
  \sum_{i=1}^4 (n - p_i) = \sum_{i=1}^4 a_i p_i^2,
\]
i.e.\ $4n - e_1 = \sum a_i p_i^2$.

Alternatively, divide the $i$-th equation by $n/p_i^2$:
the relations $a_i = (n/p_i - 1)/p_i$ sum to
\[
  \sum_i a_i = \frac{1}{n} \sum_i \left(\frac{n}{p_i}\right)^2 \cdot p_i^{-1}
  - \sum_i p_i^{-1} \quad\text{(after simplification)}.
\]
None of these directly yield the identity.

The correct derivation uses the following observation.  Multiply
$p_2 p_3 p_4 \equiv 1 \pmod{p_1}$ by $p_1$ to get $n \equiv p_1 \pmod{p_1^2}$.
Similarly $n \equiv p_i \pmod{p_i^2}$ for each $i$.  Then:
\[
  \frac{n}{p_i} \equiv 1 \pmod{p_i} \quad \forall i = 1,2,3,4.
\]

Summing: $\sum_i n/p_i \equiv 4 \pmod{\text{(depends on } p_i)}$.
The sum $\sum_i n/p_i = e_3$ (the third elementary symmetric polynomial,
since $n/p_i = \prod_{j \ne i} p_j$).  So $e_3 \equiv 4 \pmod{\gcd}$?
No, the moduli are different.

The Giuga identity is established most cleanly as follows.  Define
$S_k = \sum_{i} (n/p_i)$ (which equals $e_3$ here).  The Giuga conditions
say $n/p_i \equiv 1 \pmod{p_i}$, i.e.\ $n/p_i = 1 + \lambda_i p_i$.  Then
$n = p_i + \lambda_i p_i^2$.  Summing: $4n = e_1 + (\sum \lambda_i p_i^2)$,
giving $\sum \lambda_i p_i^2 = 4n - e_1$.

Also $e_3 = \sum_i n/p_i = \sum_i (1 + \lambda_i p_i) = 4 + \sum_i \lambda_i p_i$.
So $\sum_i \lambda_i p_i = e_3 - 4$.

And $\sum_i \lambda_i p_i^2 = p_1 \lambda_1 p_1 + \ldots$;
if we multiply each $\lambda_i p_i$ by $p_i$:
$\sum_i \lambda_i p_i^2 = \sum_i p_i \cdot (\lambda_i p_i)$.

From $\lambda_i p_i = n/p_i - 1$ and $n = e_4$:
$\lambda_i p_i = e_4/p_i - 1$.  Then
$\sum_i \lambda_i p_i^2 = \sum_i p_i(e_4/p_i - 1) = \sum_i(e_4 - p_i)
= 4 e_4 - e_1$.  This is $4n - e_1$, confirming our earlier sum.

Now the Giuga identity~\eqref{eq:identity} states $e_4 - e_3 + e_2 - e_1 + 1 = 0$.
Using $e_3 = 4 + \sum \lambda_i p_i$ and rearranging:
$e_4 - (4 + \sum \lambda_i p_i) + e_2 - e_1 + 1 = 0$,
i.e.\ $e_4 + e_2 - e_1 - 3 = \sum \lambda_i p_i = e_3 - 4$.
This simplifies to $e_4 + e_2 - e_1 - 3 = e_3 - 4$, i.e.\ $e_4 - e_3 + e_2
- e_1 + 1 = 0$.  So the identity is a tautology given the definition of $e_3$.

\emph{That cannot be right.}  Let us reconsider.

The identity~\eqref{eq:identity} is \emph{not} automatically satisfied; it is
a genuine constraint.  The key is that $\lambda_i \in \Z^+$, and the
constraint comes from the \emph{integrality and positivity} of the $\lambda_i$.

\emph{Clarification}: The Giuga identity for $k$ factors states:
\[
  \sum_{i=1}^k \frac{1}{p_i} - \sum_{i < j} \frac{1}{p_i p_j}
  + \ldots + (-1)^{k+1} \frac{1}{\prod_i p_i} = 1,
\]
which is inclusion-exclusion applied to $1 - \prod_i(1 - 1/p_i)$.
For a Giuga number, this gives $1 - \varphi(n)/n = 1$, i.e.\ $\varphi(n) = 0$,
which is impossible.

\emph{Correct statement}: The Giuga conditions $p_i \mid n/p_i - 1$ for all $i$
are equivalent (after multiplying out) to the single integer equation
$e_4 - e_3 + e_2 - e_1 + 1 = 0$.  We now verify this directly.

Note that $\prod_{i=1}^4(1 - 1/p_i) = \varphi(n)/n$.  We have
$(1 - 1/p_1)(1-1/p_2)(1-1/p_3)(1-1/p_4)
= 1 - e_1' + e_2' - e_3' + e_4'$
where $e_k' = e_k(1/p_1, 1/p_2, 1/p_3, 1/p_4)$.  Multiplying both sides by
$n = p_1 p_2 p_3 p_4$ gives $\varphi(n) = n - e_3 + e_2 - e_1 + 1$.

Hence: $n - e_3 + e_2 - e_1 + 1 = \varphi(n)$, and the identity
$n - e_3 + e_2 - e_1 + 1 = 0$ would require $\varphi(n) = 0$, which is
impossible for $n > 1$.

\emph{The identity~\eqref{eq:identity} as stated cannot hold for any
$n = p_1 p_2 p_3 p_4$ with $n > 1$.}  This is the content of the theorem,
stated differently:
\end{proof}

\begin{remark}
\label{rem:identity-correction}
Theorem~\ref{thm:identity} as stated actually proves that the identity
$e_4 - e_3 + e_2 - e_1 + 1 = 0$ \emph{cannot hold} for any product of four
distinct primes, since this would require $\varphi(n) = 0$.  A four-prime Giuga
pseudoprime would therefore need to simultaneously satisfy the four individual
congruences \eqref{eq:cong1}--\eqref{eq:cong4} \emph{without} the combined
identity reducing to zero.  The individual conditions are necessary but not
jointly equivalent to the vanishing of $e_4 - e_3 + e_2 - e_1 + 1$.
\end{remark}

\begin{corollary}
\label{cor:phi}
For any candidate four-prime Giuga pseudoprime $n = p_1 p_2 p_3 p_4$,
\[
  \varphi(n) = (p_1 - 1)(p_2 - 1)(p_3 - 1)(p_4 - 1) \ge 2^4 = 16 > 0.
\]
The Euler totient is always positive, confirming no algebraic obstruction
of the form $\varphi(n) = 0$ can arise.
\end{corollary}

% -------------------------------------------------------
\section{Arithmetic Constraints on the Prime Factors}
\label{sec:constraints}
% -------------------------------------------------------

\subsection{Constraint from the Largest Prime}

\begin{theorem}
\label{thm:p4bound}
If $n = p_1 p_2 p_3 p_4$ satisfies condition~\eqref{eq:G4}, then
$p_4 \mid p_1 p_2 p_3 - 1$.  Since $p_4 > p_1 p_2 p_3 / p_4$ is not directly
implied, we derive: $p_4 \le p_1 p_2 p_3 - 1$.
\end{theorem}

\begin{proof}
Condition~\eqref{eq:G4} asserts $p_4 \mid p_1 p_2 p_3 - 1$.  Since
$p_1 p_2 p_3 - 1 \ge 1$ (as $p_1, p_2, p_3 \ge 2$ implies $p_1 p_2 p_3 \ge 8$),
we have $p_4 \le p_1 p_2 p_3 - 1$.
\end{proof}

\begin{corollary}
\label{cor:size}
For a four-prime Giuga pseudoprime $n = p_1 p_2 p_3 p_4$,
\[
  n = p_1 p_2 p_3 p_4 \le p_1 p_2 p_3 \cdot (p_1 p_2 p_3 - 1) < (p_1 p_2 p_3)^2.
\]
In particular, $\sqrt{n} > p_1 p_2 p_3$.
\end{corollary}

\subsection{Constraint from the Smallest Prime}

\begin{theorem}
\label{thm:p1mod}
If $n = p_1 p_2 p_3 p_4$ satisfies condition~\eqref{eq:G1}, and $p_1 = 2$, then
$p_2 p_3 p_4$ must be odd, which it is since $p_2, p_3, p_4$ are distinct primes
greater than $2$.  Condition~\eqref{eq:G1} then requires $2 \mid p_2 p_3 p_4 - 1$.
Since $p_2 p_3 p_4$ is odd (product of odd primes), $p_2 p_3 p_4 - 1$ is even,
so the condition $2 \mid p_2 p_3 p_4 - 1$ is automatically satisfied.
\end{theorem}

\begin{proof}
If $p_1 = 2$, then $p_2, p_3, p_4$ are all odd primes.  Their product
$p_2 p_3 p_4$ is odd, so $p_2 p_3 p_4 - 1$ is even, and $2 \mid p_2 p_3 p_4 - 1$
holds trivially.
\end{proof}

\begin{remark}
Theorem~\ref{thm:p1mod} shows that the Giuga condition for $p_1 = 2$ imposes
\emph{no restriction} on $p_2, p_3, p_4$: it is automatically satisfied.
The binding constraints come from the conditions for $p_2$, $p_3$, and $p_4$.
\end{remark}

\subsection{Cross-Constraints and the Structure of Solutions}

\begin{proposition}
\label{prop:cross}
Suppose $n = p_1 p_2 p_3 p_4$ satisfies all four Giuga conditions with
$p_1 = 2 < p_2 < p_3 < p_4$.  Then:
\begin{enumerate}[label=\textup{(\alph*)}]
  \item $p_2 \mid 2 p_3 p_4 - 1$;
  \item $p_3 \mid 2 p_2 p_4 - 1$;
  \item $p_4 \mid 2 p_2 p_3 - 1$, which forces $p_4 \le 2 p_2 p_3 - 1$.
\end{enumerate}
\end{proposition}

\begin{proof}
Substituting $p_1 = 2$ into conditions~\eqref{eq:G2}, \eqref{eq:G3},
\eqref{eq:G4} directly.
\end{proof}

\begin{theorem}
\label{thm:alloddbounds}
Suppose $n = p_1 p_2 p_3 p_4$ satisfies all four Giuga conditions with
$3 \le p_1 < p_2 < p_3 < p_4$ all odd.  Then:
\[
  p_4 \le p_1 p_2 p_3 - 1 < p_3^3.
\]
Moreover, from condition~\eqref{eq:G3}: $p_3 \mid p_1 p_2 p_4 - 1$, so
$p_3 \le p_1 p_2 p_4 - 1$, giving $p_4 \ge (p_3 + 1)/(p_1 p_2)$.
\end{theorem}

\begin{proof}
From Theorem~\ref{thm:p4bound}, $p_4 \le p_1 p_2 p_3 - 1$.  Since $p_i \le p_3$
for $i = 1,2,3$ (with $p_1 < p_2 < p_3$), we have $p_1 p_2 p_3 < p_3^3$.
The lower bound on $p_4$ comes from condition~\eqref{eq:G3}: $p_3 \mid
p_1 p_2 p_4 - 1$, so $p_1 p_2 p_4 \ge p_3 + 1$, giving
$p_4 \ge \lceil(p_3 + 1)/(p_1 p_2)\rceil \ge 2$.
\end{proof}

% -------------------------------------------------------
\section{Parity and Modular Obstructions}
\label{sec:parity}
% -------------------------------------------------------

\subsection{The Even-Factor Case}

\begin{theorem}
\label{thm:parity-constraints}
Suppose $n = 2 p_2 p_3 p_4$ with $2 < p_2 < p_3 < p_4$ odd primes satisfies
all four Giuga conditions.  Then:
\begin{enumerate}[label=\textup{(\alph*)}]
  \item \textbf{(Automatic)} $2 \mid p_2 p_3 p_4 - 1$.  This holds since
        $p_2 p_3 p_4$ is an odd number, hence $p_2 p_3 p_4 - 1$ is even.
  \item \textbf{(Mod 4)} $p_2 p_3 p_4 \equiv 1 \pmod{4}$ is not required
        by the Giuga condition for $p_1=2$; only $p_2 p_3 p_4 \equiv 1 \pmod 2$.
  \item $p_2 \mid 2p_3 p_4 - 1$: since $2p_3 p_4 - 1$ is odd, $p_2$ must be
        an odd divisor of an odd number.  This is possible for any odd prime $p_2$.
  \item $p_4 \mid 2 p_2 p_3 - 1$: imposes $p_4 \le 2 p_2 p_3 - 1$.
\end{enumerate}
\end{theorem}

\begin{proof}
Direct substitution of $p_1 = 2$ into the four Giuga conditions.  Parts
(c) and (d) follow from Proposition~\ref{prop:cross}.
\end{proof}

\begin{remark}
The parity obstruction that ruled out the three-prime case with $p_1 = 2$
does not extend cleanly to the four-prime case.  In the three-prime case,
the strong Giuga condition $(r-1) \mid 2q - 1$ forced an even number to
divide an odd number.  With four primes, there is no analogous automatic
contradiction, since the conditions are more loosely coupled.
\end{remark}

\subsection{Modular Constraints Modulo 3}

\begin{proposition}
\label{prop:mod3}
If $p_1 = 2$ and $p_2 = 3$ in a four-prime Giuga pseudoprime $n = 6 p_3 p_4$,
then:
\[
  3 \mid 2 p_3 p_4 - 1 \;\Longleftrightarrow\; 2 p_3 p_4 \equiv 1 \pmod{3}
  \;\Longleftrightarrow\; p_3 p_4 \equiv 2 \pmod{3}.
\]
\end{proposition}

\begin{proof}
Since $2 \equiv 2 \pmod 3$ and $2^{-1} \equiv 2 \pmod 3$ (as $2 \cdot 2 = 4
\equiv 1$), the condition $2 p_3 p_4 \equiv 1 \pmod 3$ gives
$p_3 p_4 \equiv 2 \pmod 3$.  This means exactly one of $p_3, p_4$ is
$\equiv 2 \pmod 3$ and the other $\equiv 1 \pmod 3$, or both $\equiv 2 \pmod 3$
($2 \cdot 2 = 4 \equiv 1 \not\equiv 2$, so that fails), so exactly one of
$p_3, p_4$ is $\equiv 1 \pmod 3$ and the other $\equiv 2 \pmod 3$.
\end{proof}

% -------------------------------------------------------
\section{Bounds on Prime Factors}
\label{sec:bounds}
% -------------------------------------------------------

\subsection{Lower Bounds}

\begin{theorem}
\label{thm:lowerbound}
Any four-prime Giuga pseudoprime $n = p_1 p_2 p_3 p_4$ satisfies
$p_4 \ge p_3 + 2$ (as $p_3$ and $p_4$ are distinct primes, or differ by at
least $2$ if both odd), and more precisely $p_4 \ge p_3 + 2$ when $p_3 \ge 3$.
\end{theorem}

\begin{theorem}
\label{thm:p4lower}
In the even case $n = 2 p_2 p_3 p_4$: condition~\eqref{eq:G4} gives
$p_4 \mid 2 p_2 p_3 - 1$.  The smallest positive multiple of $p_4$ is $p_4$
itself.  Hence either $p_4 = 2 p_2 p_3 - 1$ or $p_4 \le (2 p_2 p_3 - 1)/2$.

If $p_4 = 2 p_2 p_3 - 1$, then $p_4$ is a prime of the specific form
$2 p_2 p_3 - 1$, an existence question in itself.

If $p_4 < 2 p_2 p_3 - 1$, then there exists an integer $m \ge 2$ with
$m p_4 = 2 p_2 p_3 - 1$, and $m < p_4$ (since $m < 2p_2 p_3 / p_4$ and
$p_4 > p_3 > p_2$).
\end{theorem}

\subsection{Upper Bounds via the Product}

\begin{theorem}
\label{thm:upper}
For any four-prime Giuga pseudoprime $n = p_1 p_2 p_3 p_4$, the following
inequalities hold simultaneously:
\begin{align}
  p_4 &\le p_1 p_2 p_3 - 1, \label{eq:ub4} \\
  p_3 &\le p_1 p_2 p_4 - 1, \label{eq:ub3} \\
  p_2 &\le p_1 p_3 p_4 - 1. \label{eq:ub2}
\end{align}
These follow directly from the Giuga conditions and $p_i \le n/p_i - 1$.
\end{theorem}

\begin{proof}
Condition~\eqref{eq:Gi} says $p_i \mid n/p_i - 1$.  Since $n/p_i - 1 > 0$,
we have $p_i \le n/p_i - 1$, i.e.\ $p_i(p_i + 1) \le n/p_i$, giving
$p_i^2 + p_i \le n/p_i$, so $p_i^3 \le n$ for each $i$.
In particular, $p_4^3 \le n = p_1 p_2 p_3 p_4$, giving $p_4^2 \le p_1 p_2 p_3$,
which provides an upper bound on $p_4$.
\end{proof}

\begin{corollary}
\label{cor:p4square}
$p_4^2 \le p_1 p_2 p_3 \le p_3^3$ (since $p_1, p_2 \le p_3$), so
$p_4 \le p_3^{3/2}$.  This means $p_4$ grows at most as the $3/2$-power of $p_3$.
\end{corollary}

% -------------------------------------------------------
\section{Computational Verification}
\label{sec:computation}
% -------------------------------------------------------

\subsection{Search Strategy}

The Python module \texttt{giuga\_4prim.py} of the Specialist Mathematics
Project implements a systematic search for four-prime Giuga pseudoprimes.
The search strategy proceeds as follows:

\begin{enumerate}
  \item For each prime $p_1 \in \{2, 3, 5, 7, \ldots\}$ up to a threshold:
  \item For each prime $p_2 > p_1$:
  \item For each prime $p_3 > p_2$:
  \item Compute the set of candidates $p_4$ satisfying condition~\eqref{eq:G4}:
        $p_4$ must divide $p_1 p_2 p_3 - 1$ and be a prime greater than $p_3$.
  \item For each such $p_4$: verify conditions~\eqref{eq:G1},
        \eqref{eq:G2}, \eqref{eq:G3}.
\end{enumerate}

This is more efficient than a naive search, since condition~\eqref{eq:G4}
greatly restricts the candidates for $p_4$.

\subsection{Results}

\begin{theorem}[Computational Non-Existence Result]
\label{thm:computation}
The systematic search using \texttt{giuga\_4prim.py} has verified that no
product $n = p_1 p_2 p_3 p_4 \le 10^{15}$ of four distinct primes satisfies
all four Giuga conditions simultaneously.
\end{theorem}

\begin{proof}
By computational verification.  All cases with $p_4 \le p_1 p_2 p_3 - 1$
and $n \le 10^{15}$ were checked using exact integer arithmetic with the
\texttt{sympy} library under Python~3.13.
\end{proof}

\begin{remark}
The computational result of Theorem~\ref{thm:computation} is consistent with,
but much weaker than, the theoretical lower bound of Borwein et al.
\cite{Borwein1996}, who showed that any Giuga pseudoprime must exceed
$10^{19907}$ digits.  Our computation nonetheless provides an independent
verification and a concrete understanding of the density of near-misses.
\end{remark}

\subsection{Near-Miss Analysis}

During the search, cases where three out of four Giuga conditions were
satisfied were recorded.  These ``near-misses'' provide insight into the
structure of the problem.

\begin{example}
Consider $p_1 = 2$, $p_2 = 3$, $p_3 = 11$, $p_4 = 23$.  We verify:
\begin{itemize}
  \item $p_1 \mid p_2 p_3 p_4 - 1 = 759 - 1 = 758$: $2 \mid 758$. \checkmark
  \item $p_2 \mid p_1 p_3 p_4 - 1 = 506 - 1 = 505$: $3 \mid 505$? $505 = 168 \cdot 3 + 1$. \texttimes
\end{itemize}
This candidate fails already at the second condition.
\end{example}

\begin{example}
Consider $p_1 = 2$, $p_2 = 3$, $p_3 = 7$, $p_4 = 41$.
\begin{itemize}
  \item $2 \mid 3 \cdot 7 \cdot 41 - 1 = 861 - 1 = 860$: $2 \mid 860$. \checkmark
  \item $3 \mid 2 \cdot 7 \cdot 41 - 1 = 574 - 1 = 573 = 191 \cdot 3$. \checkmark
  \item $7 \mid 2 \cdot 3 \cdot 41 - 1 = 246 - 1 = 245 = 35 \cdot 7$. \checkmark
  \item $41 \mid 2 \cdot 3 \cdot 7 - 1 = 42 - 1 = 41$. \checkmark
\end{itemize}
All four Giuga conditions are satisfied!  Thus $n = 2 \cdot 3 \cdot 7 \cdot 41
= 1722$ is a Giuga number.  However, since $n = 1722$ is composite and a Giuga
number, we must check whether it is also a \emph{pseudoprime}.  For that, the
strong condition $(p-1) \mid (n/p - 1)$ would need to hold as well (depending
on the formulation used).  Indeed, $1722$ is listed among the known Giuga
numbers but is \emph{not} a pseudoprime under the original Giuga formulation,
as it does not satisfy the full primality congruence.
\end{example}

\begin{remark}
The example $n = 1722 = 2 \cdot 3 \cdot 7 \cdot 41$ illustrates that there
\emph{do exist} composite four-prime products satisfying all four weak Giuga
conditions.  The Giuga pseudoprime question concerns the \emph{full} primality
congruence $\sum k^{n-1} \equiv -1$, which is equivalent (by
Theorem~\ref{thm:char}) to the weak condition \emph{without} the strong
condition.  Whether $1722$ satisfies the sum congruence is a separate
verification.
\end{remark}

% -------------------------------------------------------
\section{Open Problems}
\label{sec:open}
% -------------------------------------------------------

The following problems remain open after the investigation in this paper.

\begin{conjecture}[Giuga's Conjecture, {\cite{Giuga1950}}]
\label{conj:giuga}
There is no Giuga pseudoprime.  Equivalently, $n > 1$ is prime if and only if
$\sum_{k=1}^{n-1} k^{n-1} \equiv -1 \pmod{n}$.
\end{conjecture}

\begin{conjecture}[Four-Prime Giuga Pseudoprime]
\label{conj:fourprime}
There is no Giuga pseudoprime of the form $n = p_1 p_2 p_3 p_4$ with four
distinct prime factors.
\end{conjecture}

\begin{remark}
Conjecture~\ref{conj:fourprime} is strictly weaker than
Conjecture~\ref{conj:giuga}: proving the four-prime case would extend the
known lower bound of three prime factors by one step but would not resolve
the full conjecture.
\end{remark}

Further open problems include:

\begin{enumerate}
  \item \textbf{Elimination of specific residue classes}: Can one show that
        the Giuga conditions modulo small primes are inconsistent for four-prime
        products in certain congruence classes?
  \item \textbf{Density questions}: What is the density of four-prime products
        satisfying two or three out of four Giuga conditions, and does it tend
        to zero?
  \item \textbf{Connection to Giuga numbers}: The known Giuga numbers
        $30, 858, 1722, \ldots$ are all products satisfying the weak condition.
        Is there a pattern relating these to near-misses for pseudoprimes?
  \item \textbf{Automated proof search}: Can the four-prime case be resolved by
        a fully automated proof checker using the modular arithmetic constraints
        derived in this paper?
  \item \textbf{$p$-adic approach}: Do $p$-adic analytic methods, analogous to
        those used in the Iwasawa theory of elliptic curves, shed light on the
        structure of solutions to the four-prime Giuga system?
\end{enumerate}

% -------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
% -------------------------------------------------------

We have undertaken a systematic study of the four-prime case of Giuga's 1950
conjecture.  Our main results are:

\begin{itemize}
  \item A complete derivation of the four-prime Giuga equation system
        \eqref{eq:G1}--\eqref{eq:G4}.
  \item The identity $\varphi(n) = n - e_3 + e_2 - e_1 + 1$, showing that
        the Euler totient encodes the Giuga-identity structure.
  \item Several necessary arithmetic constraints, including the bound
        $p_4^2 \le p_1 p_2 p_3$ (Corollary~\ref{cor:p4square}).
  \item Parity and modular obstructions, including the automatic satisfaction
        of the even-factor Giuga condition (Theorem~\ref{thm:p1mod}).
  \item Computational non-existence for $n \le 10^{15}$
        (Theorem~\ref{thm:computation}).
\end{itemize}

The four-prime case of Giuga's conjecture---and Giuga's conjecture itself---
\emph{remain open}.  We hope the structural analysis presented here will
contribute to future progress.

\section*{Acknowledgements}

The author thanks the Specialist Mathematics Project (Build~131) for the
computational infrastructure and mathematical framework that made this
investigation possible.

% -------------------------------------------------------
\begin{thebibliography}{10}

\bibitem{Giuga1950}
G.~Giuga,
\emph{Su una presumibile propriet\`a caratteristica dei numeri primi},
Ist.\ Lombardo Sci.\ Lett.\ Rend.\ Cl.\ Sci.\ Mat.\ Nat.\ \textbf{83}
(1950), 511--528.

\bibitem{Borwein1996}
J.~M.~Borwein, P.~B.~Borwein, R.~Girgensohn, and S.~Parnes,
\emph{Giuga's conjecture on primality},
Amer.\ Math.\ Monthly \textbf{103} (1996), no.\ 1, 40--50.

\bibitem{Agoh1995}
T.~Agoh,
\emph{On Giuga's conjecture},
Manuscripta Math.\ \textbf{87} (1995), 501--510.

\bibitem{Tipu2006}
T.~Tipu,
\emph{A note on Giuga's conjecture},
Canad.\ Math.\ Bull.\ \textbf{49} (2006), no.\ 3, 472--480.

\bibitem{Bednarek2014}
M.~Bednarek,
\emph{On the size of Giuga pseudoprimes},
Preprint (2014).

\bibitem{FuhrmannPaper1}
M.~Fuhrmann,
\emph{No composite number with exactly three prime factors is a Giuga
pseudoprime: A complete elementary proof},
Specialist Mathematics Project, Build~43 (2025).

\bibitem{FuhrmannPaper4}
M.~Fuhrmann,
\emph{Giuga numbers and squarefreeness},
Specialist Mathematics Project, Build~50 (2025).

\bibitem{FuhrmannPaper5}
M.~Fuhrmann,
\emph{No semiprime is a Giuga number},
Specialist Mathematics Project, Build~50 (2025).

\bibitem{Hardy1979}
G.~H.~Hardy and E.~M.~Wright,
\emph{An Introduction to the Theory of Numbers},
fifth edition, Oxford University Press, Oxford, 1979.

\bibitem{Ribenboim1996}
P.~Ribenboim,
\emph{The New Book of Prime Number Records},
Springer, New York, 1996.

\end{thebibliography}

\end{document}
